### 2014-04-17 docs/seminar/casestudy.tex
more changes to all sections

\subsection{The Platform}\label{sec:platform}
Essential components of the platform as well as the work-flow of searching for the translators are depicted in Figure \ref{fig:architecture}. The platform consist of four main components: \textit{Ranking}, \textit{Proficiency Estimator}, \textit{Scheduler} and \textit{Profiles}.

\begin{figure*}
\begin{center}
\includegraphics[width=12cm]{figures/dataflow.png}
\caption{Translator-Expert Search Work Flow
\label{fig:architecture}}
\end{center}
\end{figure*}

The \textit{Profiles} stores translator profile information i.e. source and target languages, offered price and translation duration per word, and the number of translations the translator has performed for the same client. The \textit{Scheduler} system figures out the delivery time based on timetable of translators. The scheduler builds an efficient data structure to calculate the delivery time in a reasonable response time. The details of the process are out of scope of the paper. The \textit{Proficiency Estimator} system stores the previously-translated documents and indexes them using Lucene. The similarity between query and indexed documents is used as a basis for the estimation of translator's proficiency. In fact, the proficiency value is achieved by aggregating the documents similarity scores. The applied aggregation function is described in Section\ref{sec:apply}. At last, the \textit{Ranking} system uses all the data provided in the previous steps to create the ranking model. The \textit{Ranking} system applies Learning to Rank techniques to return the most related translators to the client. The training data is provided by a group of annotators who are familiar with the business of company using an evaluation system. The evaluation system suggests three translators and the annotators rate them by comparing between their factors. In order to prevent bias in evaluation, the translators are suggested randomly and without name and picture. Applied learning to rank methods and results are described in Section\ref{sec:apply}.

The client submits a document and searches for translators with a specific target language. Based on the query document, the framework figures out the offered price, delivery time, proficiency of translators as well as the number of cooperation times related to each translator. The ranking system processes the calculated values and offers the most related translators to the client.

>>>>

\subsection{The Platform}\label{sec:platform}
Essential components of the platform as well as the workflow of searching for the translators are depicted in Figure \ref{fig:architecture}. The platform consists of four main components: \textit{Ranking}, \textit{Proficiency Estimator}, \textit{Scheduler} and \textit{Profiles}.

\begin{figure*}
\begin{center}
\includegraphics[width=12cm]{figures/dataflow.png}
\caption{Translator-Expert Search Workflow
\label{fig:architecture}}
\end{center}
\end{figure*}

The \textit{Profiles} component stores translator profile information i.e. source and target languages, offered price and translation duration per word, as well as the number of translations the translator has performed for the same client. The \textit{Scheduler} system calculates the delivery time based on the timetable of each translator. The scheduler builds an efficient data structure to calculate the delivery time in a reasonable response time. The details of the process are out of scope of the paper. In this report we focus on the Proficiency Estimator and the Ranking elements.

The \textit{Proficiency Estimator} sub-system stores the previously-translated documents of each translator and indexes them using Lucene. The similarity between query and indexed documents is used as a basis for the estimation of translator's proficiency for the task at hand. The proficiency score is obtained by aggregating the documents' similarity scores. Different aggregation functions are analyzed in Section\ref{sec:apply}. Finally, the \textit{Ranking} sub-system uses all the data generated in the previous steps to create the ranking model. It uses Learning to Rank techniques to return the most relevant candidate translators. The training data is provided by a group of annotators  familiar with the business of the company, using an evaluation system created specifically for this generating this ground truth. The evaluation system presents three translators and the annotators rank them based on the values of each of their attributes (i.e. proficiency, delivery time, price, cooperation). In order to prevent bias in evaluation, the translators are suggested randomly and without name and picture. The applied learning to rank methods and their results are described in Section\ref{sec:apply}.

%The client submits a document and searches for translators with a specific target language. Based on the query document, the framework figures out the offered price, delivery time, proficiency of translators as well as the number of cooperation times related to each translator. The ranking system processes the calculated values and offers the most related translators to the client.

### 2014-04-17 docs/seminar/evaluation.tex
more changes to all sections

\section{Experimental Results}
\label{sec:apply}
In this section, we applied different approaches on the platform. By comparing the methods, we aim to discover the most appropriate one regarding to the project's characteristics and data.

\subsection{Aggregation Functions}
Translator's proficiency is one of the attributes used for creating the ranking model. In order to achieve the value, we aggregate the similarity values of the previous translated documents with the query document. Selecting the most appropriate aggregation function is the objective of the current section.

Similar to \cite{agg-gp2}, we compare three aggregation algorithms: $GP2$, $Top5$ and $Top1$. $Top1$ and $Top5$ are two common forms of $TopN$ aggregation algorithm which refers to algorithm that summarizes the $N$ top documents (i.e. $Top1$ only using the top associated document to rank the candidates).

In order to evaluate aggregation functions results, we use the assessment of the proof-readers after every translation as the basis for the comparison. **the reason**. Therefore, the more similar the results of the algorithm to the assessments are the better the algorithm is.

By calculating the value of each mentioned algorithm on all finished translations, we achieve three lists of translations each annotated with an additional aggregation value field. In the next step, in order to compare the algorithms, we test the correlation of each list with the list of translations annotated with proof-readers' assessments.

In order to calculate the correlation value, we applied Spearman Rank Order and Kendall Rank Correlation as two common methods. Table \ref{table:correlation} shows the results of Spearman correlation coefficient ($r_s$) and Kendall's tau coefficient ($\tau$) using $181$ records of purchased orders. In addition, it represents the Significance Test of both methods. For Spearman's test, p-values are computed using algorithm AS 89 \cite{as89}.

As it is usual in many cases, the coefficient value of Spearman is slightly higher than Kendall's. The outcome shows an approximately weak correlation between aggregation functions and feedbacks of proof-readers. Regarding to P-Value of significance test shown in Table\ref{table:correlation}, a meaningful relation between GP2 and feedbacks can be considered. Top1 which has the worst values in the table shows a meaningless and near random correlation though.

@@

\subsection{Learning To Rank}
In order to accumulate data required for ranking model, we conduct a survey filled by eight human annotators. The questions of survey represent three translators each with four criteria (price, delivery time, proficiency and number of cooperation times). As it is mentioned before, in order to prevent bias in results, the name and the picture of the translators are hidden from the annotators. The annotators rate the questions from one to three based on \textit{Total Order} strategy. The accumulated data consists of $400$ annotated list and overall $1200$ records.

We use the RankLib library\footnote{http://sourceforge.net/p/lemur/wiki/RankLib} to apply Learning to Rank methods. The library provides the  implementation of some Learning to Rank algorithms as well as evaluation measures in Java. By splitting the data in train, validation and test datasets, we use 5-Fold Cross-Validation on each run. One pointwise, three pairwise and two listwise methods are applied.

Since evaluation of the results should be applied on whole the result list (with $3$ items), we have to use relevance grading evaluation measures like DCG, NDCG or ERR. In the current study, NDGC and ERR both with rank position at $3$ are used.

Because of the short final predicted list ($3$ items), every evaluator measure returns high result value even for a random ranker. This issue urges the need to run the evaluation on random data in order to find a baseline for comparison. We tackle the problem with two approaches. In the first approach, we run all the algorithms on data with random generated labels. In order to increase the accuracy of results, the process of generating randomized label is repeated $5$ times. The second approach is developing a simple ranker which randomly predicts the rank of each record. In fact, the first tests whether there is anything to learn in the data and the second examines if the learning algorithms learn from the data.

>>>>

\section{Experimental Results}
\label{sec:apply}
In this section, we applied the different approaches presented in the previous section within our platform. By comparing the methods, we aim to discover the most appropriate one considering to the project's characteristics and data.

\subsection{Aggregation Functions}
Translator's proficiency on the topic of the query is one of the attributes used for creating the ranking model. In order to obtain the value, we aggregate the similarity values of the previously translated documents with the query document. Selecting the most appropriate aggregation function is the objective of the current section.

In order to evaluate aggregation functions results, we use the assessment of the proof-readers after every translation. As mentioned before, the hypothesis is that a translator is likely to do a better job if she is familiar with the vocabulary of the query. We can then use the proof-readers assessments as a gold standard and observe which of the three aggregation methods correlates more with the scores given by the proof-readers.

Similar to \cite{agg-gp2}, we compare three aggregation algorithms: $GP2$, $Top5$ and $Top1$. $GP2$ is the state of the art, while $Top1$ and $Top5$ are two common forms of $TopN$ aggregation algorithm which refers to algorithm that summarizes the $N$ top documents. Furthermore, $Top1$ can be interpreted as the maximum similarity score, and follows the intuition that it is sufficient for a translator to have had only one highly similar translation job before in order to do a good job on the new task.

As input data we have 181 translation orders collected from the live system. For each of them, we know the translator who performed the translation (including all her previously translated documents), the text of the document to be translated, and the score given by the proof-reader. Because this is historical data, evaluating the aggregation method cannot be done by giving the same translation task to different translators. However, vy calculating the value of each mentioned algorithm on all finished translations, we achieve three lists of translations each annotated with an additional aggregation value field. Then, in order to compare the algorithms, we test the correlation of each list with the list of translations annotated with proof-readers' assessments. Again, the assumption being that if we rank translation jobs by aggregation scores, the top elements would have high proof-reader scores, while the bottom elements lower scores.

%mihai: put a plot here. On the X-axis the proof-reader scores and on the Y-axis the aggregation scores.

In order to calculate the correlation value, we applied Spearman Rank Order and Kendall Rank Correlation as two common methods. Table \ref{table:correlation} shows the results of Spearman correlation coefficient ($r_s$) and Kendall's tau coefficient ($\tau$) using the 181 records of purchased orders. In addition, it represents the Significance Test of both methods. For Spearman's test, p-values are computed using algorithm AS 89 \cite{as89}.

As it is usual in many cases, the coefficient value of Spearman is slightly higher than Kendall's. The outcome shows an approximately weak correlation between aggregation functions and feedbacks of proof-readers. Regarding to p-Value of significance test shown in Table\ref{table:correlation}, a meaningful relation between GP2 and feedbacks can be considered. Top1 which has the worst values in the table shows a meaningless and near random correlation though.

@@

\subsection{Learning To Rank}
In order to accumulate data required for ranking model, we conduct a survey with eight human annotators. The questions of survey represent three translators each with four criteria (price, delivery time, proficiency and number of cooperation times). As mentioned before, in order to prevent bias in results, the name and the picture of the translators are hidden from the annotators. The annotators rate the questions from one to three based on \textit{Total Order} strategy. The accumulated data consists of $400$ annotated list and overall $1200$ records.

We use the RankLib library\footnote{http://sourceforge.net/p/lemur/wiki/RankLib} to apply a large set of Learning to Rank methods. The library provides the  implementation of some Learning to Rank algorithms as well as evaluation measures in Java. By splitting the data in train, validation and test datasets, we use 5-Fold Cross-Validation on each run. One pointwise, three pairwise and two listwise methods are tested.

Since evaluation of the results should be applied on  the entire result list (with $3$ items), we have to use relevance grading evaluation measures like DCG, NDCG or ERR. In the current study, NDGC and ERR both with rank position at $3$ are used.

Because of the short final predicted list ($3$ items), every  measure returns a very high score. In order to understand the real benefit of our framework compared with just presenting the un-ranked set of three translators, we evaluate two random approaches. In the first approach, we run all the algorithms on data with random generated labels. In order to increase the accuracy of results, the process of generating randomized label is repeated $5$ times. The second approach is developing a simple ranker which randomly predicts the rank of each record. In fact, the first tests whether there is anything to learn in the data and the second examines if the learning algorithms learn from the data.

### 2014-04-17 docs/seminar/introduction.tex
more changes to all sections

This paper addresses the problem of searching translators as experts. We offer a novel translator-expert retrieval platform and evaluate different expert retrieval methods based on the practical implementation of the platform. In contrast to common expert retrieval systems, we include significant factors of searching a translator (such as price and delivery time) in the automated decision process as well as the relevance of the documents each translator has previously translated to the query document. The proposed method has two distinct components: A proficiency estimation phase, in which different aggregation algorithms related to documents of translators have been studied and compared; and a Learning-to-Rank phase in which different features are tested under all state of the art Learning-to-Rank methods based on a manually created ground truth, which we make available together with the source code of our application under GPL\footnote{\url{http://github/anonymized}}.

>>>>

This paper addresses the problem of searching translators as experts. We offer a novel translator-expert retrieval platform and evaluate different expert retrieval methods based on the practical implementation of the platform. In contrast to common expert retrieval systems, we include significant factors of searching a translator (such as price and delivery time) in the automated decision process as well as the relevance of the documents each translator has previously translated to the query document. The proposed method has two distinct components: A proficiency estimation phase, in which different aggregation algorithms related to documents of translators have been studied and compared; and a Learning-to-Rank phase in which different features are tested under all state of the art Learning-to-Rank methods based on a manually created ground truth, which we make available together with the source code of our application under GPL\footnote{\url{https://github.com/neds/expert-retrieval-translators}}.

### 2014-04-17 docs/seminar/methods.tex
more changes to all sections

\cite{trec2005} propose two principal approaches  in the expertise retrieval area based on probabilistic language modeling techniques. They were formalized as so-called \textit{candidate models} and \textit{document models}. The candidate-based approach, also referred to as profile-based method, builds a textual representation of candidate experts and then ranks them based on the query. The document models first find documents relevant to the topic and then locate the experts associated with these documents \cite{er}.

@@

As well as ranking techniques, aggregation functions have a significant effect on the performance of expert retrieval systems. Aggregate tasks are those where documents' similarities are not
the final outcome, but instead an intermediary component. In expert search, a ranking of candidate persons with relevant expertise to a query is generated after aggregation of their related documents \cite{agg-learning}.
\subsection{Aggregation Functions}
The aggregation function has a significant impact on the performance of Expert Retrieval system. As a usual scenario in expert retrieval systems, first each document related to an expert is scored and ranked regarding to query. Then, the top $N$ document scores associated with a candidate expert are aggregated in order to rank the experts.

The effect of different features on the aggregation function is studied in \cite{agg-gp2}. They show that the number of documents is an important factor, such that the performance of different queries are optimal for different values of $N$. Comparing query-based features using statistical measures, they infer that the document features (such as TF, IDF) may not, in general, be able to predict the optimal number of documents to aggregate for each query. In contrast, individual Expert Features have been shown to be more informative such that relevant experts are associated with a higher ranked document than non-relevant experts. More interestingly, relevant experts are associated with less documents on average.

\cite{agg-vote} looks to expert search as a voting problem, where documents vote for the candidates with relevant expertise. Eleven data fusion methods as well as three statistically different document weighting models were tested. In practice, the approach considers both the number of documents and expert features regarding to the ranking score of the documents. The results show that while some of adapted voting techniques most likely outperform others, the proposed approach is effective when using the appropriate one.

Later on, focusing on related features discussed before \cite{agg-gp2} introduces a new aggregation method. It uses genetic programming to learn a formula for the weights of document associations within the candidate profiles. The formula denoted as $GP2$ is as follows:

@@

\subsection{Learning To Rank}
Learning to rank refers to machine learning techniques for training the model in a ranking task. Due to importance of ranking problems, learning to rank has been drawing broad attention in the machine learning community recently.

In the learning to rank approach, the ranking problem is transformed to classification, regression and ordinal classification, and existing methods and techniques for solving machine learning problems are applied. As \cite{l2r-intro} points out, the relation between learning to rank and ordinal classification is that, in ranking, one cares more about accurate ordering of objects, while in ordinal classification, one cares more about accurate ordered-categorization of objects.

One primitive step in accumulating data required for learning to rank, is relevance judgment normally done by human annotators. \cite{l2r-book} introduces three main strategies as follows.
\begin{itemize}
\item \textit{Relevance degree}: In this method, the annotator specifies whether an object is relevant or not to the query. It can be either in binary judgment or by specifying the degree of relevance (e.g., Perfect, Excellent, Good, Fair, or Bad).
\item \textit{Pairwise preference}: The annotator compares a pair of objects in order to specify which one is more relevant with regards to a query.
\item \textit{Total order}: The annotator specifies the total order of all objects with respect to a query by rating each object.
\end{itemize}

@@

In the pointwise approach, the ranking problem is transformed to classification, regression or ordinal classification. Therefore, the group structure of ranking is ignored in this approach \cite{l2r-intro}. Linear or polynomial regression are widely used methods regarding to the approach.

The pairwise approach transforms the ranking problem into pairwise classification or regression. In fact, it cares about the relative order between two documents. Similar to the pointwise, the pairwise approach also ignores the group structure of ranking\cite{l2r-intro}. Here is a brief explanation of some pairwise algorithms:

@@

\cite{l2r-book} compares the algorithms by applying on different data-sets. It concludes that listwise techniques are in general the most effective among the others. However, the choice of the learning evaluation measure and the rank cutoff may have a noticeable impact on the effectiveness of the learned model \cite{l2r-when}.

@@

Recently, \cite{l2r-err} have proposed a new evaluation metrics called Expected Reciprocal Rank (ERR) which claims to model user's satisfaction with search results better than the DCG metric. \cite{l2r-err} addresses the underlying independence assumption of DCG that a document in a given position has always the same gain and discount independently of the documents shown above it. It asserts that based on research on modeling user click behavior \cite{l2r-clickmodel1,l2r-clickmodel2}, the likelihood a user examines the document at rank $i$ is dependent on how satisfied the user was with previously observed documents in the ranked list. In other words, it assumes that a user is more likely to stop browsing if they have already seen one or more highly relevant documents. Introducing the ERR formula, \cite{l2r-err} claim that results reflect real user browsing behavior better and quantifies user satisfaction more accurately than DCG.

>>>>

Cao and colleagues~\cite{trec2005} propose two principal approaches  in the expertise retrieval area based on probabilistic language modeling techniques. They were formalized as so-called \textit{candidate models} and \textit{document models}. The candidate-based approach, also referred to as profile-based method, builds a textual representation of candidate experts and then ranks them based on the query. The document models first find documents relevant to the topic and then locate the experts associated with these documents \cite{er}.

In either of the two models, aggregation functions have a significant effect on the performance of expert retrieval systems. Aggregate tasks are those where documents' similarities are not
the final outcome, but instead an intermediary component. In expert search, a ranking of candidate persons with relevant expertise to a query is generated after aggregation of their related documents \cite{agg-learning}.

@@

For the task at hand, we found that the two methods have to be used in two different steps. Aggregation can be used to bring together in one value elements essentially of the same nature. In this case - query similarity scores of different documents. The test collection at hand relies for its topical similarity exclusively on term frequencies, as there are no hyperlinks, metadata, or other sources of information in the documents. As a second step, in order to bring in attributes orthogonal to topical similarity, learning to rank methods are an obvious choice. In the following, we describe related work related to these two aspects, and in doing so prepare the ground for our experiments, which we present in the next section.

\subsection{Aggregation Functions}
The aggregation function has a significant impact on the performance of Expert Retrieval system~\cite{agg-gp2}. As a usual scenario in expert retrieval systems, first each document related to an expert is scored and ranked regarding to query. Then, the top $N$ document scores associated with a candidate expert are aggregated in order to rank the experts.

MacDonald and Ounis \cite{agg-vote} consider expert search as a voting problem, where documents vote for the candidates with relevant expertise. Eleven data fusion methods as well as three statistically different document weighting models were tested in their experiments. In practice, the approach considers both the number of documents and expert features regarding to the ranking score of the documents. The results show that while some of adapted voting techniques most likely outperform others, the proposed approach is effective when using the appropriate one.

Cummins and colleagues~\cite{agg-gp2}  study the effect of different features on the aggregation function. They show that the number of documents is an important factor, in that the performance of different queries are optimal for different values of $N$. Comparing query-based features using statistical measures, they infer that the document features (such as TF, IDF) may not, in general, be able to predict the optimal number of documents to aggregate for each query. In contrast, individual Expert Features have been shown to be more informative such that relevant experts are associated with a higher ranked document than non-relevant experts. More interestingly, relevant experts are associated with less documents on average.

Focusing on these features Cummins et al. \cite{agg-gp2} introduce a new aggregation method. It uses genetic programming to learn a formula for the weights of document associations within the candidate profiles. The formula, denoted as $GP2$, is as follows:

@@

\subsection{Learning To Rank}
Learning to rank refers to machine learning techniques for training a model in a ranking task. Due to importance of ranking problems, learning to rank has been drawing broad attention in the machine learning community recently.

In the learning to rank approach, the ranking problem is transformed to classification, regression and ordinal classification, and existing methods and techniques for solving machine learning problems are applied. As Hang\cite{l2r-intro} points out, the relation between learning to rank and ordinal classification is that, in ranking, one cares more about accurate ordering of objects, while in ordinal classification, one cares more about accurate ordered-categorization of objects.

One primitive step in accumulating data required for learning to rank, is relevance judgments, normally done by human annotators. Lie \cite{l2r-book} presents the three main strategies in learning to rank:
\begin{itemize}
\item \textit{Relevance degree}: In this method, the annotator specifies whether an object is relevant or not to the query. It can be either in binary judgment or by specifying the degree of relevance (e.g., Perfect, Excellent, Good, Fair, or Bad).
\item \textit{Pairwise preference}: The annotator compares a pair of objects in order to specify which one is more relevant with regards to a query.
\item \textit{Total order}: The annotator specifies the total order of all objects with respect to a query by rating each object.
\end{itemize}

@@

In the pointwise approach, the ranking problem is transformed to classification, regression or ordinal classification. Therefore, the group structure of ranking is ignored in this approach \cite{l2r-intro}. Here, linear or polynomial regression are widely used methods.

The pairwise approach transforms the ranking problem into pairwise classification or regression. In fact, it cares about the relative order between two documents. Similar to the pointwise approach, the pairwise method also ignores the group structure of ranking~\cite{l2r-intro}. Here is a brief explanation of some pairwise algorithms:

@@

Lie \cite{l2r-book} compares the algorithms by applying on different data-sets. It concludes that listwise techniques are in general the most effective among the others. However, the choice of the learning evaluation measure and the rank cutoff may have a noticeable impact on the effectiveness of the learned model \cite{l2r-when}.

\subsection{Evaluation}

@@

Recently, Chapelle and Zhang~\cite{l2r-err} have proposed Expected Reciprocal Rank (ERR) which claims to model user's satisfaction with search results better than the DCG metric. Their work addresses the underlying independence assumption of DCG that a document in a given position has always the same gain and discount independently of the documents shown above it. It asserts that based on research on modeling user click behavior \cite{l2r-clickmodel1,l2r-clickmodel2}, the likelihood a user examines the document at rank $i$ is dependent on how satisfied the user was with previously observed documents in the ranked list. In other words, it assumes that a user is more likely to stop browsing if they have already seen one or more highly relevant documents. Introducing the ERR formula, Chapelle and Zhang claim that results reflect real user browsing behavior better and quantifies user satisfaction more accurately than DCG.
